---
title: "TP1"
output: html_notebook
---

Question 0 :

```{r}
library(dplyr)
health <- read.table("health.txt",header=TRUE, sep=" ", dec=".")
health <- health[2:5]
print(health)
```

Question 1 :

```{r}
library(dplyr)
library(corrplot)
correlation_matrix<-cor(health)
corrplot(correlation_matrix, order = 'hclust',addrect = 3)
```


Question 2 :

```{r}
y <-lm(y~.,data=health)

library(car)
library(ggfortify)
autoplot(y,1)         #P1 assumption
```
We can see that the blue line isautoplot(y,1)  not a line horizontal and close to 0. So assumption P1 is not satisfied.

```{r}
autoplot(y,3)         #P2 assumption
```
We can see that the blue line is not aligned around 1. So assumption P1 is not satisfied.

```{r}
set.seed(0)
durbinWatsonTest(y)   #P3 assumption 
```

We can see that the p-value is greater than 5%, so we don't reject H0. So assumption P3 is satisfied.

```{r}
autoplot(y,2)       #P4 assumption 
```

We can see that the points fall approximately along a strait line (y=x) but we have some points that are far from this line. So assumption P4 is not satisfied.

Question 3:
```{r}
library(GGally)
ggpairs(health, progress = F)
```
We can see that there is a quadratic relation with age in function of y, age = sqrt(y).

```{r}
health2 <- health
health2$agesqrt <- health$age^2
health2 <- health2[1:24,]

y2 <- lm(y~.,data=health2)

autoplot(y2,1)
autoplot(y2,3)
durbinWatsonTest(y2)
autoplot(y2,2) 
```

