---
title: "TP 2"
output:
  html_document: default
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Exercice 2

Let $X_1, \ldots, X_n$ be i.i.d. random variables following the exponential distribution $\mathcal{E}(\lambda)$, i.e. $\mathbb{E} [X_1] = \lambda^{-1}$.

## 1

### (a) 

Simulate 10000 realizations of the distribution $\mathcal{E}(\lambda)$ for $\lambda = 2$ using the inverse function method.

### (b) 

Check with a histogram and a Quantile-Quantile plot that the distribution of this sample matches the distribution $\mathcal{E}(\lambda)$.


```{r}
n <- 10000
lam <- 2

f_exp <-function(n,lam){
  return(-log(runif(n)) / lam)
}

X <- f_exp(n,lam)
```

```{r}
Y <- rexp(n,lam)
x <- seq(0,10,0.1)

par(mfrow = c(1,2))

hist(X,freq = FALSE, ylim = c(0, 1.5))
lines(density(X), col = 'blue')
curve(dexp(x, lam),from = 0, add = TRUE, col = "red")

qqplot(X,Y, main = "QQ-Plot between real Exp(2) and estimated one")
```

## 2

Recall that $S_n = X_1 + . . . + X_n$ follows the gamma distribution $\Gamma(n, \lambda)$, i.e. $mathbb{E}[S_n] = n\lambda−1$.

### (a) 

Based on this result, simulate 10000 realizations of the gamma distribution $\Gamma(n, \lambda)$ with $\lambda = 1.5$ and $n = 10$.

### (b) 

Graphically verify that the distribution of this sample matches the gamma distribution $\Gamma(n, \lambda)$.

```{r}
n <- 10
lam <- 1.5
N <- 10000

f_gamma <- function(n,lam,N){
  return(rowSums(matrix(f_exp(n*N,lam),N,n)))
}

S <- f_gamma(n,lam,N)
```

```{r}
Y <- rgamma(N, shape = n, rate = lam)
x <- seq(0, 50, 0.1)

par(mfrow = c(1,2))

hist(S,freq = FALSE)
lines(density(S), col = 'blue')
curve(dgamma(x, n, lam), add = TRUE, col = "red")

qqplot(S,Y,main = "QQ-Plot between real Gamma(10,1.5) and estimated one")
```

## 3

Let $N = sup \{ n \ge 1 : S_n \le 1 \}$ (by convention $N = 0$ if $S_1 > 1$). Then $N$ follows the Poisson distribution $\mathcal{P}(\lambda)$. 

### (a) 

Based on this result, simulate 10000 realizations of the Poisson distribution $\mathcal{P}(\lambda)$ with $\lambda = 4$.

### (b) 

Graphically verify that the distribution of this sample matches the Poisson distribution $\mathcal{P}(\lambda)$.


```{r}
n <- 10000
lam <- 4

f_poiss <- function(n, lam) {
  res <- numeric(n)
  for (i in 1:n) {
    sum_exp <- 0
    k <- 0
    while (sum_exp <= 1) {
      sum_exp <- sum_exp + rexp(1, lam)
      k <- k + 1
    }
    res[i] <- k - 1
  }
  return(res)
}

P <- table(f_poiss(n,lam)) / n
Y <- table(rpois(n,lam)) / n
mat <- rbind(P, Y)

barplot(mat, main = "Frequency of P", xlab = "value of P", ylab = "frequency",
col = c(rgb(1,0,0,1/2), rgb(0,0,1,1/2)), beside = TRUE,
legend.text = c("approximation of Poisson", "Poisson ref"))

```
