---
title: "TP3"
output: html_notebook
---
```{r}
library(dplyr)
library(rmarkdown)
library(MASS)
library(ggplot2)
library(GGally)
library(ggfortify)
library(car)
```


```{r}
#load the data 
ozone <- read.csv("ozone.txt",header = T, sep = "", dec = "." )
ozone$vent <- as.factor(ozone$vent)
ozone$temps <- as.factor(ozone$temps)
```

```{r}
#simple linear model maxO3 against T12
mod <- lm(maxO3 ~ T12,data = ozone )
summary(mod)
```

```{r}
#simple linear regression 
autoplot(mod,1:4)
durbinWatsonTest(mod)
```

By the simple linear regression, we have the assumption :
  P[1] is validated as the blue line centered on 0
  P[2] is validated as the blue line centered on 1
  P[3] not validated as the Durbin Watson test p-value lower than 5%
  P[4] is validated as QQ-plot aligned with y=x line 
  
The model is not well define

```{r}
#ANOVA model 
mod_vent <- lm(maxO3 ~ vent , data = ozone) #maxO3 against vent
mod_temps <- lm(maxO3 ~ temps, data = ozone)  #maxO3 against temps
mod_both <- lm(maxO3 ~ temps * vent , data = ozone)  #maxO3 against temps and vent
```

```{r}
summary(mod_temps)
```

```{r}
#ANOVA regression
autoplot(mod_temps,1:4)
durbinWatsonTest(mod_temps)
```
We have that non of the assumption is validates.


```{r}
summary(mod_vent)
```

```{r}
#ANOVA regression
autoplot(mod_vent,1:4)
durbinWatsonTest(mod_vent)
```
We have that non of the assumption is validates.

```{r}
summary(mod_both)
```

```{r}
#ANOVA regression
autoplot(mod_both,1:4)
durbinWatsonTest(mod_both)
```
We have that non of the assumption is validates.


```{r}
#multiple linear model maxO3 against all co variables
mod_all <- lm(maxO3 ~., data = ozone)
summary(mod_all)
```

```{r}
#multiple linear regression
autoplot(mod_all,1:4)
durbinWatsonTest(mod_all)
```

By the simple linear regression, we have the assumption :
  P[1] is validated as the blue line centered on 0
  P[2] is validated as the blue line centered on 1
  P[3] is validated as the Durbin Watson test p-value lower than 5%
  P[4] is validated as QQ-plot aligned with y=x line 
  
The model is not well define, but there is to many variables, we will reduce it.

```{r}
#graph of the possible correlation between co variables
ggpairs(ozone, progress = F)
```

```{r}
#model Forward by step AIC
modForw <- stepAIC(mod_all, maxO3 ~ maxO3 + T9 + T12 + T15 + Ne9 + Ne12 + Ne15 + Vx9 + Vx12 + Vx15 + maxO3v + vent + temps,trace=F, direction=c("forward"))
summary(modForw)
```
It is the same model as the model with all variables.

```{r}
#model Backward by step AIC
modBack <- stepAIC(mod_all, maxO3 ~. ,trace=F, direction=c("backward"))
summary(modBack)
```

This method select the model maxO3 against T12,Ne9,Vx9 and maxO3v.
```{r}
#model Both ward by step AIC
modBoth <- stepAIC(mod_all, maxO3 ~ maxO3 + T9 + T12 + T15 + Ne9 + Ne12 + Ne15 + Vx9 + Vx12 + Vx15 + maxO3v + vent + temps ,trace=F, direction=c("both"))
summary(modBoth)
```

It is the same model as the backward model.

```{r}
#Test the better model
AIC(modForw,modBack,modBoth)
```

The better model is the backward model as it has the lower AIC.

```{r}
#implement new data
new <- data.frame(T12=18, Ne9=3, Vx9=0.7, maxO3v=85)

#test if the model well predict it 
predict(modBack, new, interval = "confidence")
```
The prediction fit with the interval of confidence, the model well predict data.

